---
title: "Code-along, Week 01: Emmy nominees"
author: "Alex Homer"
date: "23 September 2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(wesanderson) #For colour scales later
```

*(Before running code chunks from this file, go to Session > Set Working Directory > To Files Pane Location in RStudio.)*

*(Also, as you'll see at the end, there's a data collection error happened with this dataset that affects all of the analysis below.  You have been warned!)*

## Read data
The data are ^[People disagree on whether you should say "The data is" or "The data are".  I've had it drummed into me that "data" is a plural word---singular "datum" or "data point"---but you are free to disagree with me!] drawn from the "Tidy Tuesday" project: [2021 Week 39: Emmy Awards and Nominees](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-09-21/readme.md) (credit: [Emmys website](https://www.emmys.com/awards/nominations/award-search)/[Susie Lu](https://www.susielu.com/data-viz/emmy-2017)).  They have been lightly cleaned.

```{r read-data, message = FALSE}
nominees_original <- readRDS("data/nominees_clean.rds")
```

*(Some of the code used in the subsequent sections is relatively advanced!  Don't worry if you don't understand what every line is doing.  We'll be covering all of these by the end of the course.)*

## Cleanup
We need to do a little more clean-up first, because some shows are multiple-listed when multiple people shared the same award.

```{r cleanup, message = FALSE}
nominees <- nominees_original %>%
  select(category, type, title, distributor, year, page, page_id) %>%
  distinct()
```

## Who won the most Emmys?
```{r winner, message = FALSE}
winners <- nominees %>%
  filter(type == "Winner") %>%
  select(-type)

(
  winners_count <- winners %>%
    count(title, sort = TRUE)
)
```

*Saturday Night Live* won the most Emmys.  Let's see a graph!

```{r winner-plot, message = FALSE}
#First make a new data frame with an "other" category.
cutoff <- 15

winners_count %>% filter(n >= cutoff) %>%
  mutate(title = fct_reorder(title, n)) %>%
  ggplot(aes(x = n, y = title)) +
  geom_col() +
  labs(
    x = "Number of wins",
    y = "Show",
    title = "Shows with most Emmy wins",
    subtitle = paste("Shows with at least", cutoff, "wins shown"),
    caption = "Source: Emmys website/Susie Lu"
  ) +
  theme_minimal()
```

What if we wanted to see which distributor was responsible for each winning show?  We'll need to go back to the original data, since we've already "thrown away" that information in the `winners_count` data frame.

```{r distribs, message = FALSE}
cbbPalette <- c(
  "#000000", "#E69F00", "#56B4E9", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7"
)

winners %>% count(title, distributor) %>%
  filter(n >= cutoff) %>%
  mutate(title = fct_reorder(title, n)) %>%
  ggplot(aes(x = n, y = title, fill = distributor)) +
  geom_col() +
  labs(
    x = "Number of wins",
    y = "Show",
    fill = "Distributor",
    title = "Shows with most Emmy wins",
    subtitle = paste("Shows with at least", cutoff, "wins shown"),
    caption = "Source: Emmys website/Susie Lu"
  ) +
  scale_fill_manual(values = cbbPalette) +
  theme_minimal()
```

The colour palette used is taken from [The R Cookbook](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette) and is colourblind-friendly.

## HBO v Netflix
HBO looks to have had quite a few of the most successful shows (in terms of Emmy nominations, at least), but one of the stories people chose to tell with these data on Tidy Tuesday was the rise of Netflix.  Let's see if we can graph that.

```{r hbo-v-netflix, message = FALSE}
hbo_netflix <- nominees %>%
  filter(distributor == "HBO" | distributor == "Netflix")

#We need to account for the fact that winners were also nominees
hbo_netflix_extra <- hbo_netflix %>%
  filter(type == "Winner") %>%
  mutate(type = "Nominee")

hbo_netflix_year <- hbo_netflix %>%
  bind_rows(hbo_netflix_extra) %>%
  count(distributor, type, year)

hbo_netflix_year %>%
  arrange(year) %>%
  mutate(distributor = fct_reorder(distributor, -n, last)) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line(aes(colour = distributor, linetype = type), size = 2) +
  scale_colour_manual(values = wes_palette("IsleofDogs1")) +
  labs(
    x = "Year",
    y = "Number of nominations",
    colour = "Distributor",
    title = "Netflix overtook HBO for Emmy nominations in 2020...",
    subtitle = "... and for winners in 2021",
    caption = "Source: Emmys website/Susie Lu"
  ) +
  theme_minimal()
```

Maybe the different line types are confusing...  We could fix that by faceting, but we'll do that another time.

The `wes_palette` function used here is from the [`wesanderson`](https://cran.r-project.org/web/packages/wesanderson/index.html) package, and provides colourblind-friendly colour palettes inspired by the films of [Wes Anderson](https://en.wikipedia.org/wiki/Wes_Anderson) (who is known for "deliberately limited colour palette[s]" ^[Source: [Wikipedia](https://en.wikipedia.org/wiki/Wes_Anderson#Visual_style)] in his movies).

## What happened in 2014?
See the dip in 2014?  Let's investigate that.

```{r 2014, message = FALSE}
winners %>% count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(
    x = "Year",
    y = "Total number of winners",
    title = "The number of Emmy winners rose sharply in the 21st Century...",
    subtitle = "... apart from in 2014, apparently",
    caption = "Source: Emmys website/Susie Lu"
  ) +
  theme_minimal()
```

Let's investigate...

```{r investigate}
nominees_2014 <- nominees_original %>%
  filter(year == 2014)

nominees_2013 <- nominees_original %>%
  filter(year == 2013)

nominees_2014 %>% select(category) %>%
  distinct()
```

The 2014 dataset only appear to contain acting awards!  Whereas for 2013, if we look at awards without the words "Actor" or "Actress" in the title.

```{r investigate-further}
nominees_2013 %>% select(category) %>%
  filter(!str_detect(category, "Actor")) %>%
  filter(!str_detect(category, "Actress")) %>%
  distinct()
```

We see lots of different awards.  This appears to be a data problem---it looks like the 2014 awards haven't been scraped properly from the website.

```{r breaking-bad}
nominees %>% filter(
  title == "Breaking Bad",
  category == "Outstanding Drama Series"
)
```

Yep, *Breaking Bad* [won this award in 2014](https://www.emmys.com/awards/nominations/award-search?search_api_views_fulltext=&field_is_winner=1&field_award_category%5B%5D=44541&field_celebrity_details_field_display_name=&field_show_details_field_nominee_show_nr_title=&field_nominations_year_op=%3E%3D&field_nominations_year%5Bvalue%5D=2014-01-01&field_nominations_year_1_op=%3C%3D&field_nominations_year_1%5Bvalue%5D=2014-01-01).  There's been a data collection error, and of course that affects our whole analysis!

*(This wasn't building up to a deliberate conclusion---I realised these things in the order presented here.  Beware: this could happen to you!  At some point I will go back and re-scrape the data including the 2014 shows properly, and re-do the plots.)*